---
title: "Extracting from Worldclim2"
date: "7/17/20"
author: "Max Frissell"
output:
  prettydoc::html_pretty:
    theme: cayman
---

```{r Packages, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

library(dplyr)
library(tidyr)
library(magrittr)
library(ggplot2)
library(here)
library(hexbin)

# Note raster masks 'tidyr::extract' and 'dplyr::select'
library(raster)
```

```{r SRDBSetup, include = FALSE}
## Selects the lines I want to use from the SRDB

# Read in entire SRDB and pull only desired columns, removing rows with missing values
read.csv(here("Data", "srdb-data.csv")) %>%
  dplyr::select(Site_ID, Latitude, Longitude, Leaf_habit, MAT, MAP, Rs_annual, Rh_annual, RC_annual) %>%
  subset(!is.na(Site_ID) & !is.na(Latitude) & !is.na(Longitude) & !is.na(Leaf_habit) & !is.na(Rs_annual) & !is.na(Rh_annual) & RC_annual >= 0 & RC_annual <= 1) -> srdb
```

## Investigating Data included in the SRDB

Here are some graphs, showing how different aspects change between sites with different leaf habits.

``` {r SRDBLeafHabit}
## Makes some graphs showing how the leaf habit affects other variables

ggplot(srdb, aes(x = Leaf_habit, fill = Leaf_habit)) +
  geom_bar() +
  theme_minimal()

ggplot(srdb, aes(x = Leaf_habit, y = Rs_annual, fill = Leaf_habit)) +
  geom_violin(draw_quantiles = c(.25, .5, .75)) +
  theme_minimal()

ggplot(srdb, aes(x = Leaf_habit, y = RC_annual, fill = Leaf_habit)) +
  geom_violin(draw_quantiles = c(.25, .5, .75)) +
  theme_minimal()

ggplot(srdb, aes(x = Longitude, y = Latitude, col = Leaf_habit)) +
  geom_point() +
  theme_minimal()
```

## Pulling the Climate Data and Comparing the Distribution to that in the SRDB

The first is a plot of the global distribution of mean annual temperature (MAT) and precipitation (MAP). 
The second also includes the values from all SRDB sites, plotted on top of the first graph.

```{r PullingFromWC2}
## This chunk pulls climate data from WorldClim2 and makes some graphs from the mean annual temperature and precipitation data

# Download worldclim data for precip and tmean if necessary, into w10/ folder
precip <- getData("worldclim", path = here(), var = "prec", res = 10, download = !file.exists("wc10/prec1.hdr"))
tmean <- getData("worldclim", path = here(), var = "tmean", res = 10, download = !file.exists("wc10/wc10/tmean1.hdr"))

# Pull out cosore dataset latitudes and longitudes
srdb %>%
  dplyr::select(Site_ID, Longitude, Latitude) -> srdb_coords

# MAP data that matches the srdb coordinates
raster::extract(precip, srdb_coords[2:3]) -> precip_coords
apply(precip_coords, 1, sum) -> MAP_WC
cbind(srdb_coords, MAP_WC) -> map_coords

# The same for MAT
raster::extract(tmean, srdb_coords[2:3]) -> tmean_vals
apply(tmean_vals, 1, mean) -> MAT_WC
cbind(map_coords, MAT_WC)  %>%
  # Temp data is stored in degC * 10, so we need to divide to get back to degC
  mutate(MAT_WC = MAT_WC / 10) -> srdb_points

# Extract global climate space data
raster::as.data.frame(precip, xy = TRUE) %>%
  drop_na() -> precip_global

# Calculate annual sum for precip...
precip_global %>%
  dplyr::select(-x, -y) %>%
  apply(1, sum) -> map_global

raster::as.data.frame(tmean, xy = TRUE) %>%
  drop_na() -> tmean_global

# ...and mean for temperature
tmean_global %>%
  dplyr::select(-x, -y) %>%
  apply(1, mean) -> mat_global

# Create tibble with corresponding coordinates
tibble(x = tmean_global$x, y = tmean_global$y, mat = as.vector(mat_global)) -> mat
tibble(x = precip_global$x, y = precip_global$y, map = as.vector(map_global)) -> map

left_join(map, mat, by = c("x", "y")) %>%
  # Temp data is stored in degC * 10, so we need to divide to get back to degC
  mutate(mat = mat / 10) -> map_mat_global

# Just the global distribution of MAT and MAP
p <- ggplot() +
  geom_hex(data = map_mat_global,
             aes(x = mat, y = map), bins = 100, na.rm = TRUE) +
  scale_fill_viridis_c(name = "Grid cells", begin = 0.85, end = 0) +
  theme_minimal() +
  labs(x = "MAT (째C)", y = "MAP (mm)")

# With points from the SRDB sites
withPoints <- ggplot() +
  geom_hex(data = map_mat_global,
             aes(x = mat, y = map), bins = 100, na.rm = TRUE) +
  scale_fill_viridis_c(name = "Grid cells", begin = 0.85, end = 0) +
  geom_point(data = srdb_points, aes(x = MAT_WC, y = MAP_WC),
             color = "black", shape = 1, size = 1.5, na.rm = TRUE) +
  theme_minimal() +
  labs(x = "MAT (째C)", y = "MAP (mm)")

print(p)
print(withPoints)

MAT_WC <- MAT_WC/10
# Add the worldclim MAT and MAP values to the end of the srdb data
allData = cbind(srdb, MAT_WC, MAP_WC)
```

## Comparing the WorldClim Climate Data to the Reported Data in the SRDB

This is a graph comparing the mean annual temperature (MAT) reported by whoever made the entry in the SRDB to the MAT predicted by WorldClim2 for those coordinates.

``` {r ComparingMAT}
## Makes a graph comparing the mean annual temperatures in the SRDB and in their corresponding WorldClim2 values

attach(allData)

ggplot(allData, aes(x = MAT_WC, y = MAT)) + 
  geom_point(size = 1.5, color = "chartreuse3") +
  theme_minimal() +
  labs(title = "Comparing WorldClim2 MATs to SRDB MATs", x = "WorldClim2 MAT (째C)", y = "Reported SRDB MAT (째C)") +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", size = 1)
```

Now the same with mean annual precipitation (MAP).

``` {r ComparingMAP}
## Makes a graph comparing the mean annual precipitations in the SRDB and in their corresponding WorldClim2 values

attach(allData)

ggplot(allData, aes(x = MAP_WC, y = MAP)) + 
  geom_point(size = 1.5, color = "chartreuse3") +
  theme_minimal() +
  labs(title = "Comparing WorldClim2 MAPs to SRDB MAPs", x = "WorldClim2 MAP (mm)", y = "Reported SRDB MAP (mm)") +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", size = 1)
```

Here are some of the relationships between mean annual temperature and precipitation and Rs and Rh.

``` {r WCRespirationScatterplots}
## Makes some scatterplots with the SRDB data

ggplot(allData, aes(x = MAT_WC, y = Rs_annual)) +
  geom_point(color = "chartreuse3") +
  ylim(0, 3000) + 
  theme_minimal()

ggplot(allData, aes(x = MAT_WC, y = Rh_annual)) +
  geom_point(color = "chartreuse3") +
  ylim(0, 2000) +
  theme_minimal()

ggplot(allData, aes(x = MAP_WC, y = Rs_annual)) +
  geom_point(color = "chartreuse3") +
  ylim(0, 3000) + 
  theme_minimal()

ggplot(allData, aes(x = MAP_WC, y = Rh_annual)) +
  geom_point(color = "chartreuse3") +
  ylim(0, 2000) +
  theme_minimal()
```

Now, here are how the climates of sites with different leaf habits change.

``` {r LeafHabitAndClimate}
ggplot(allData, aes(x = Leaf_habit, y = MAT_WC, fill = Leaf_habit)) +
  geom_violin(draw_quantiles = c(.25, .5, .75)) +
  theme_minimal()

ggplot(allData, aes(x = Leaf_habit, y = MAP_WC, fill = Leaf_habit)) +
  geom_boxplot() +
  theme_minimal()

ggplot(allData, aes(x = MAT_WC, y = MAP_WC, col = Leaf_habit)) +
  geom_point() +
  theme_minimal()
```

## Pulling Mycorrhizae Data and Investigating it

``` {r ExtractingMyco}
## Extracts data on mycorrhizae for each SRDB point from global dataset from Soudzilovskaia et al. 2019
## Note: the .TIF files used here are not included in the repo and were downloaded from https://github.com/nasoudzilovskaia/Soudzilovskaia_NatureComm_MycoMaps/tree/master/Maps_Myco_veg_current on 7/14/20

# Get all of the global data for each type of myco (again, files are pre-downloaded and not in the GitHub repo)
am = raster(here("Data", "MycDistrAM_current.TIF"))
em = raster(here("Data", "MycDistrEM_current.TIF"))
er = raster(here("Data", "MycDistrER_current.TIF"))
nm = raster(here("Data", "MycDistrNM_current.TIF"))

# Extract the myco data for each coordinate pair in the SRDB database and add it to the rest of the data
AM_percent = raster::extract(am, allData[, c(3, 2)])
EM_percent = raster::extract(em, allData[, c(3, 2)])
ER_percent = raster::extract(er, allData[, c(3, 2)])
NM_percent = raster::extract(nm, allData[, c(3, 2)])
allData = cbind(allData, AM_percent, EM_percent, ER_percent, NM_percent)

# Throw out the entries where there isn't myco data
allData = allData[!is.na(allData$AM_percent),]
```

These are plots of the percentage of plants that use difference kinds of mycorrhizae (the last being none) for each of the SRDB sites.

``` {r MycoGraphs}
## Plot the myco percentages with latitude and longitude.

ggplot(allData, aes(x = Longitude, y = Latitude, col = AM_percent)) + 
  geom_point(size = 1.5) + 
  lims(x = c(-180, 180), y = c(-50, 80), color = c(0, 100)) + 
  theme_minimal()

ggplot(allData, aes(x = Longitude, y = Latitude, col = EM_percent)) + 
  geom_point(size = 1.5) + 
  lims(x = c(-180, 180), y = c(-50, 80), color = c(0, 100)) + 
  theme_minimal()

ggplot(allData, aes(x = Longitude, y = Latitude, col = ER_percent)) + 
  geom_point(size = 1.5) + 
  lims(x = c(-180, 180), y = c(-50, 80), color = c(0, 100)) + 
  theme_minimal()

ggplot(allData, aes(x = Longitude, y = Latitude, col = NM_percent)) + 
  geom_point(size = 1.5) + 
  lims(x = c(-180, 180), y = c(-50, 80), color = c(0, 100)) + 
  theme_minimal()
```

This is a comparison of how the abundance of the different mycorrhizae types changes with mean annual temperature.

``` {r Myco&MAT}
## Plot the myco percentages vs MAT

ggplot(allData, aes(x = MAT_WC, y = AM_percent)) + 
  geom_point(size = 1.5, color = "chartreuse3") + 
  theme_minimal()

ggplot(allData, aes(x = MAT_WC, y = EM_percent)) + 
  geom_point(size = 1.5, color = "chartreuse3") + 
  theme_minimal()

ggplot(allData, aes(x = MAT_WC, y = ER_percent)) + 
  geom_point(size = 1.5, color = "chartreuse3") + 
  theme_minimal()

ggplot(allData, aes(x = MAT_WC, y = NM_percent)) + 
  geom_point(size = 1.5, color = "chartreuse3") + 
  theme_minimal()
```

As you can see, there seem to be noticeable, but weak, correlations.

None of these seem to show correlations with mean annual precipitation, with the exception of ER mycorrhizae.

``` {r Myco&MAP}
## Plot the ER myco percentage vs MAP
## ER is the only one that showed any visible correlation

ggplot(allData, aes(x = MAP_WC, y = ER_percent)) + 
  geom_point(size = 1.5, color = "chartreuse3") + 
  theme_minimal()
```

Here are the breakdowns of the the prevalence of the different mycorrhizae types in ecosystems with diffent leaf habits.

``` {r Myco&LeafHabit}
## Makes violin plots of the prevalence of the different mycorrhizae types, separated by leaf habit

ggplot(allData, aes(x = Leaf_habit, y = AM_percent, fill = Leaf_habit)) +
  geom_violin(draw_quantiles = c(.25, .5, .75)) +
  theme_minimal()

ggplot(allData, aes(x = Leaf_habit, y = EM_percent, fill = Leaf_habit)) +
  geom_violin(draw_quantiles = c(.25, .5, .75)) +
  theme_minimal()

ggplot(allData, aes(x = Leaf_habit, y = ER_percent, fill = Leaf_habit)) +
  geom_boxplot() +
  theme_minimal()

ggplot(allData, aes(x = Leaf_habit, y = NM_percent, fill = Leaf_habit)) +
  geom_violin(draw_quantiles = c(.25, .5, .75)) +
  theme_minimal()
```

## Pulling Global Biomass Data and Investigating it

``` {r ExtractingBiomass}
## Extracts data on above and below ground biomass for each SRDB point from global dataset from Spawn et al. 2020
## Note: the .TIF files used are not included in the repo and need to be downloaded from the "sethspawn/globalBiomassC" GitHub repo

raster(here("Data", "aboveground_biomass_carbon_2010.tif")) %>%
  raster::extract(allData[, c(3, 2)]) -> BM_aboveground

raster(here("Data", "belowground_biomass_carbon_2010.tif")) %>%
  raster::extract(allData[, c(3, 2)]) -> BM_belowground

# Add these biomass values to the larger dataset
allData = cbind(allData, BM_aboveground, BM_belowground)

# Remove values where there are no biomass data
allData = allData[!is.na(allData$BM_aboveground) & !is.na(allData$BM_belowground),]
```

These are maps of the above and below ground biomasses at all of the sites in the SRDB.

There is also a summary of all of the data so far.

``` {r BiomassMaps}
## Creates a map with the values of above/below ground biomas data at each SRDB site

ggplot(allData, aes(x = Longitude, y = Latitude, col = BM_aboveground)) +
  geom_point(size = 1.5) +
  theme_minimal()

ggplot(allData, aes(x = Longitude, y = Latitude, col = BM_belowground)) +
  geom_point(size = 1.5) +
  theme_minimal()
```

I searched for some correlations between the biomass values and some of the other data, like MAT, MAP, root contribution, and Rs, but I didn't really find much.
Here is everything that is somewhat interesting. (The first graph is really pretty)

``` {r BiomassRelationships}
## Shows relationships between biomass and other variables

ggplot(allData, aes(x = BM_aboveground, y = BM_belowground)) +
  geom_point(color = "chartreuse3") +
  theme_minimal()

ggplot(allData, aes(x = Leaf_habit, y = BM_aboveground, fill = Leaf_habit)) +
  geom_violin(draw_quantiles = c(.25, .5, .75)) +
  theme_minimal()

ggplot(allData, aes(x = Leaf_habit, y = BM_belowground, fill = Leaf_habit)) +
  geom_violin(draw_quantiles = c(.25, .5, .75)) +
  theme_minimal()
```