---
title: "Extracting from Worldclim2"
date: "7/7/20"
author: "Max Frissell"
output:
  prettydoc::html_pretty:
    theme: cayman
---

```{r Packages, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

library(dplyr)
library(tidyr)
library(magrittr)
library(ggplot2)
library(here)
library(hexbin)

# Note raster masks 'tidyr::extract' and 'dplyr::select'
library(raster)
```

```{r SRDBSetup, include = FALSE}
# Selects which lines will be read (NULL = not read, NA = read) and reads them in from SRDB
lines = c(rep("NULL", 10), NA, "NULL", "NULL", NA, NA, rep("NULL", 18), NA, NA, rep("NULL", 11), NA, rep("NULL", 5), NA, rep("NULL", 32))
srdb = read.csv(here("Data", "srdb-data.csv"), colClasses = lines)
```

## Pulling the Climate Data and Comparing the Distribution to that in the SRDB

The first is a plot of the global distribution of mean annual temperature (MAT) and precipitation (MAP). 
The second also includes the values from all SRDB sites, plotted on top of the first graph.

```{r PullingFromWC2}
# Download worldclim data for precip and tmean if necessary, into w10/ folder
precip <- getData("worldclim", path = here(), var = "prec", res = 10, download = !file.exists("wc10/prec1.hdr"))
tmean <- getData("worldclim", path = here(), var = "tmean", res = 10, download = !file.exists("wc10/wc10/tmean1.hdr"))

# Pull out cosore dataset latitudes and longitudes
srdb %>%
  dplyr::select(Site_ID, Longitude, Latitude) -> srdb_coords

# MAP data that matches the srdb coordinates
raster::extract(precip, srdb_coords[2:3]) -> precip_coords
apply(precip_coords, 1, sum) -> MAP_WC
cbind(srdb_coords, MAP_WC) -> map_coords

# The same for MAT
raster::extract(tmean, srdb_coords[2:3]) -> tmean_vals
apply(tmean_vals, 1, mean) -> MAT_WC
cbind(map_coords, MAT_WC)  %>%
  # Temp data is stored in degC * 10, so we need to divide to get back to degC
  mutate(MAT_WC = MAT_WC / 10) -> srdb_points

# Extract global climate space data
raster::as.data.frame(precip, xy = TRUE) %>%
  drop_na() -> precip_global

# Calculate annual sum for precip...
precip_global %>%
  dplyr::select(-x, -y) %>%
  apply(1, sum) -> map_global

raster::as.data.frame(tmean, xy = TRUE) %>%
  drop_na() -> tmean_global

# ...and mean for temperature
tmean_global %>%
  dplyr::select(-x, -y) %>%
  apply(1, mean) -> mat_global

# Create tibble with corresponding coordinates
tibble(x = tmean_global$x, y = tmean_global$y, mat = as.vector(mat_global)) -> mat
tibble(x = precip_global$x, y = precip_global$y, map = as.vector(map_global)) -> map

left_join(map, mat, by = c("x", "y")) %>%
  # Temp data is stored in degC * 10, so we need to divide to get back to degC
  mutate(mat = mat / 10) -> map_mat_global

# Just the global distribution of MAT and MAP
p <- ggplot() +
  geom_hex(data = map_mat_global,
             aes(x = mat, y = map), bins = 100, na.rm = TRUE) +
  scale_fill_viridis_c(name = "Grid cells", begin = 0.85, end = 0) +
  theme_minimal() +
  labs(x = "MAT (°C)", y = "MAP (mm)")

# With points from the SRDB sites
withPoints <- ggplot() +
  geom_hex(data = map_mat_global,
             aes(x = mat, y = map), bins = 100, na.rm = TRUE) +
  scale_fill_viridis_c(name = "Grid cells", begin = 0.85, end = 0) +
  geom_point(data = srdb_points, aes(x = MAT_WC, y = MAP_WC),
             color = "black", shape = 1, size = 1.5, na.rm = TRUE) +
  theme_minimal() +
  labs(x = "MAT (°C)", y = "MAP (mm)")

print(p)
print(withPoints)

MAT_WC <- MAT_WC/10
# Add the worldclim MAT and MAP values to the end of the srdb data
allData = cbind(srdb, MAT_WC, MAP_WC)
```

## Comparing the WorldClim Climate Data to the Reported Data in the SRDB

This is a graph comparing the mean annual temperature (MAT) reported by whoever made the entry in the SRDB to the MAT predicted by WorldClim2 for those coordinates.

``` {r ComparingMAT}
attach(allData)

ggplot(allData, aes(x = MAT_WC, y = MAT)) + 
  geom_point(size = 1.5, color = "chartreuse3") +
  theme_minimal() +
  labs(title = "Comparing WorldClim2 MATs to SRDB MATs", x = "WorldClim2 MAT (°C)", y = "Reported SRDB MAT (°C)") +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", size = 1)
```

Now the same with mean annual precipitation (MAP).

``` {r ComparingMAP}
attach(allData)

ggplot(allData, aes(x = MAP_WC, y = MAP)) + 
  geom_point(size = 1.5, color = "chartreuse3") +
  theme_minimal() +
  labs(title = "Comparing WorldClim2 MAPs to SRDB MAPs", x = "WorldClim2 MAP (°C)", y = "Reported SRDB MAP (°C)") +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", size = 1)
```
