---
title: "Making Some Graphs"
author: "Max Frissell"
date: "7/1/2020"
output: 
  html_document:
    theme: cerulean
---

``` {r ReadData, include = F}
library(here)
library(magrittr)
library(ggplot2)
library(png) #load packages for use later

path = here("Data", "srdb-data.csv")
allData = read.csv(path) 
attach(allData) #get all SRDB data and attach it for use

RCData = allData[RC_annual > -1,] #only the data with RC values
```

### Some Basic Scatterplots from the SRDB Data

#### *They're all Longitude and Latitude, so they are basically world maps.*

First, here is a plot of all values of the root contribution (RC) to soil respiration (Rs) and where they were taken.

```{r GraphRC, warning = F}
ggplot(RCData, aes(x = Longitude, y = Latitude, col = RC_annual)) + 
  geom_point(size = 2) + 
  xlim(-180, 180) + ylim(-50, 80) + 
  theme_bw()
```

This is a similar graph, but instead of showing the root contribution, it shows the total annual soil respiration in grams of carbon per square meter. 
Note, this only includes sites with Rs values less than 5,000. 
Without filtering out some of the very high values that are likely due to measurement error, the color scale doesn't show anything.

```{r GraphRs, warning = F}
ggplot(allData[Rs_annual <= 6000,], aes(x = Longitude, y = Latitude, col = Rs_annual)) + 
  geom_point(size = 2) + 
  xlim(-180, 180) + ylim(-50, 80) + 
  theme_bw()
```

This is the mean annual temperature (MAT) in degrees Celsius, vs the location.

```{r GraphMAT, warning = F}
ggplot(allData[MAT >= -20,], aes(x = Longitude, y = Latitude, col = MAT)) + 
  geom_point(size = 2) + 
  xlim(-180, 180) + ylim(-50, 80) + 
  theme_bw()
```

Similarly, this is the mean annual precipitation (MAP) in millimeters.

```{r GraphMAP, warning = F}
ggplot(allData[MAP >= 0,], aes(x = Longitude, y = Latitude, col = MAP)) + 
  geom_point(size = 2) + 
  xlim(-180, 180) + ylim(-50, 80) + 
  theme_bw()
```

### Histograms of Relevent Data

Now here is the distribution of values in the SRDB for the soil respiration root contribution.
The values that are less than 0 or greater than 1 must be wrong because these are proportions.

```{r RCHistogram, message = F}
hist(RCData$RC_annual, main = "Histogram of Annual RC's", xlab = "Annual RC", breaks = 20, col = "4") #makes the histogram
```

This is the histogram of annual soil respiration values.
There were a few very large outliers that were removed.
This kept the maximum x axis value small enough for the histogram to actually have meaning.

```{r RsHistogram, message = F}
hist(allData$Rs_annual[Rs_annual <= 5000], main = "Histogram of Annual Rs's", xlab = "Annual Rs (grams of C per square meter)", breaks = 20, col = "4") #makes the histogram
```

#### *Doesn't that just look beautiful! By far my favorite graph here*

This is a histogram of the mean annual temperature values, in degrees Celsius.

```{r MATHistogram, message = F}
hist(allData$MAT, main = "Histogram of Mean Annual Temperatures", 
     xlab = "MAT (degrees C)", breaks = 40, col = "4") #makes the histogram
```

This is a histogram of mean annual precipitation values, in millimeters.

```{r MAPHistogram, message = F}
hist(allData$MAP, main = "Histogram of Mean Annual Precipitations",
     xlab = "MAP (millimeters)", breaks = 20, col = "4") #makes the histogram
```

### Now this is just a relationship that I stumbled into and the graph looks cool

This is the relationship between the distance from the equator in degrees and mean annual temperature.
Everyone would expect their to be a relationship, but it turned out pretty well and I just like how the graph looks.

```{r LattitudeAndMAT, message = F, warning = F}
ggplot(allData[MAT >= -20,], aes(x = abs(Latitude), y = MAT)) + 
  geom_point(color = "4", size = 2) + 
  theme_bw() + 
  labs(x = "Distance from Equator (degrees)", y = "MAT (degrees C)")
```

That was fun.
